---
title: "HW2 -- Bingo"
author: "Anthony Bernardi"
date: "2023-08-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 - Bingo Pseudocode



*a ) A function that creates a bingo card.*



No default argument or input. 

Returns a 5x5 matrix, the columns labeled B, I, N, G, O.

Column 1 values randomly sampled from numbers 1 to 15.

Column 2 values randomly sampled from numbers 16 to 30.

Column 3 values randomly sampled from numbers 31 to 45, except for a "Free space" 
marked from the start of the game for all cards.

Column 4 values randomly sampled from numbers 46 to 60.

Column 5 values randomly sampled from numbers 61 to 75.


*b) Function to play bingo by marking the card.*


Two default arguments; the player's card and the number called from the hopper for all players.

Function reads in player's card matrix.

Function reads number called from the hopper, and if the numbers match, the corresponding card is marked.

If there is no match, nothing is done with the player's card.


*c) Function to read the player's bingo card and decide whether or not they've won the game.*


Input is the player's marked bingo card.

If the bingo card has 5 consecutive marks vertically, horizontally, or diagonally, print statement about how the player has won the game.

If the card does not have 5 consecutive marks, do nothing and continue with the game.


## 2 -- R Code

## 2.1 -- Function to Create the Bingo Card

```{r}

# Introducing the function with no default argument 
get_card <- function(){
  
  # generating the ranges to sample the bingo cards from 
  range_1 <- seq(from = 1 , to = 15)
  range_2 <- seq(from = 16 , to = 30)
  range_3 <- seq(from = 31, to = 45)
  range_4 <- seq(from = 46 , to = 60)
  range_5 <- seq(from = 61 , to = 75)
  
  # taking a random sample in each case that will eventually be our rows and columns 
  row_1_sample <- sample(range_1 , size = 5)
  row_2_sample <- sample(range_2 , size = 5)
  row_3_sample <- sample(range_3 , size = 5)
  row_4_sample <- sample(range_4 , size = 5)
  row_5_sample <- sample(range_5 , size = 5)
  
  # creating and generating the actual matrix  
  card <- matrix(data = c(row_1_sample , row_2_sample , row_3_sample , 
                          row_4_sample , row_5_sample) ,
                 nrow = 5 , ncol = 5)
  
  # specifying rows and column names 
  row_names <- c(1 , 2, 3, 4, 5)
  title <- c('B', 'I', 'N', 'G', 'O')
  
  dimnames(card) <- list(row_names , title)
  
  
  # ensuring everyone gets their free space 
  
  card[3,3] <- 0
  
  print(card)
  
}
```


```{r , echo = FALSE}
player_1 <- get_card()
```

## 2.2 -- Function to Play Bingo by Marking the Card

```{r}

# Introducing the function once we have the player's card and a number from the
# hopper. 

marked <- function(number, card){
  
  
  # need to loop over each row and column of a matrix 
  # looping over every element of the matrix, or player's card, will let us
  # know if there is a match 
  # matches are marked with a 0
  for (i in 1:nrow(card)){
    for (j in 1:ncol(card)){
      if (card[i,j] == number){
        card[i,j] <- 0
      } else {
      }
    }
  }
  # want to be sure the player's card updates if there is a match, or stays 
  # the same if nothing happens.
  card <- card
  print(card)
}
```

Let's use an example where the player with the above card has a match.
Say a number on the player's first row and second column is called from the
hopper. We write this in the R code and run it implicitly. We'll run both 
functions and mark that card with a '0'.

```{r , echo = FALSE}
x <- player_1[1,2]


player_1 <- marked(x , player_1)
```


## 2.3 -- Function that Considers the Card and Decides Whether or Not the Player Has Won the Game

```{r}

# function that takes in the player's current card and scans for matches 
# along the rows, columns, and diagonally 

winner <- function(card){
  
  # takes just the existing/marked card as an input and later checks for zeroes 
  
  # check the rows first 
  for (i in 1:nrow(card)){
    if(sum(card[i,]) == 0){
      print('Bingo!')
    } else{
      
    }
  }
  
  # now checking the columns 
  for (i in 1:ncol(card)){
    if(sum(card[,i]) ==0){
      print('Bingo!')
    } else{
      
    }
  }
  
  # now we can check the diagonals
  # checking if the diagonal or if the reverse of the diagonal have our 
  # five zeroes for a match/bingo 
  
  if (sum(diag(card)) == 0 | sum(rev(diag(card))) == 0){
    print('Bingo!')
  } else{
    
  }
  
  # for loop to create a diagonal in the opposite direction 
  # calling it the reverse diagonal, going from the fifth row 
  # of the first column to the fifth column of the first row 
  
  # creating a place to store the reverse diagonal 
  reverse_diag <- 0
  for (i in 1:5){
    reverse_diag[i] <- card[i , 6-i]
  }
  
  # clause to check for a bingo, have an or to ensure that both 
  # directions of the diagonal are scanned by the function
  
  if (sum(reverse_diag) == 0 | sum(rev(reverse_diag)) == 0){
    print('Bingo!')
  } else if (length(card[card == 0]) <= 5){
    print('Keep playing!')
  }
  
}

```

Do we have a winner?

The 'Winner' function will print 'Keep playing!' if the player's card doesn't have a bingo, and will print 'Bingo!' once there is. 

```{r , echo = FALSE}
winner(player_1)
```


## Extra example - A Winning Card 

For posterity, let's consider a card where the player has a Bingo and has won 
the game. We'll start with their card. 

```{r , echo = FALSE}

player_2 <- get_card()

```

We'll then replace just the first row with all 0's -- meaning that their card 
has been marked in all five spaces in the first row. If only Bingo were this 
easy!

```{r , echo = FALSE}

player_2[1,] <- rep(x = 0 ,  times = 5)

player_2

```

Finally, we'll run our 'Winner' function with this card and the player should 
get a message that they've won the game!

```{r , echo = FALSE}

winner(player_2)

```

