---
title: "HW4 - Bingo Winners"
author: "Anthony Bernardi"
date: "2023-09-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We'll start with introducing our three functions from before. 

```{r}
get_card <- function(){
  
  # generating the ranges to sample the bingo cards from 
  range_1 <- seq(from = 1 , to = 15)
  range_2 <- seq(from = 16 , to = 30)
  range_3 <- seq(from = 31, to = 45)
  range_4 <- seq(from = 46 , to = 60)
  range_5 <- seq(from = 61 , to = 75)
  
  # taking a random sample in each case that will eventually be our rows and columns 
  row_1_sample <- sample(range_1 , size = 5)
  row_2_sample <- sample(range_2 , size = 5)
  row_3_sample <- sample(range_3 , size = 5)
  row_4_sample <- sample(range_4 , size = 5)
  row_5_sample <- sample(range_5 , size = 5)
  
  # creating and generating the actual matrix  
  card <- matrix(data = c(row_1_sample , row_2_sample , row_3_sample , 
                          row_4_sample , row_5_sample) ,
                 nrow = 5 , ncol = 5)
  
  # specifying rows and column names 
  row_names <- c(1 , 2, 3, 4, 5)
  title <- c('B', 'I', 'N', 'G', 'O')
  
  dimnames(card) <- list(row_names , title)
  
  
  # ensuring everyone gets their free space 
  
  card[3,3] <- 0
  
  print(card)
  
}
```

And calling them in each case. 

```{r , echo = FALSE}
card <- get_card()
```

```{r}
marked <- function(number, card){
  
  
  # need to loop over each row and column of a matrix 
  # looping over every element of the matrix, or player's card, will let us
  # know if there is a match 
  # matches are marked with a 0
  for (i in 1:nrow(card)){
    for (j in 1:ncol(card)){
      if (card[i,j] == number){
        card[i,j] <- 0
      } else {
      }
    }
  }
  # want to be sure the player's card updates if there is a match, or stays 
  # the same if nothing happens.
  card <- card
  #move_counter <- move_counter + 1
  #print(paste('Number of Moves:', move_counter))
  print(card)
}
```

Sampling a random number to call from the Hopper. 
```{r}
n <- sample(1:75 , 1)

card <- marked(n, card)


```


Finally, our winner function. 

```{r}
winner <- function(card){
  
  # takes just the existing/marked card as an input and later checks for zeroes 
  
  # check the rows first 
  for (i in 1:nrow(card)){
    if(sum(card[i,])  == 0){
      print('Bingo! -- By Rows!')
      return(card)
    } else{
      
    }
  }
  
  # now checking the columns 
  for (i in 1:ncol(card)){
    if(sum(card[,i])  == 0){
      print('Bingo! -- By Columns!')
      return(card)
    } else{
      
    }
  }
  
  # now we can check the diagonals
  # checking if the diagonal or if the reverse of the diagonal have our 
  # five zeroes for a match/bingo 
  
  if (sum(diag(card)) == 0 | sum(rev(diag(card))) == 0){
    print('Bingo! -- Diagonal!')
    return(card)
  } else{
    
  }
  
  # for loop to create a diagonal in the opposite direction 
  # calling it the reverse diagonal, going from the fifth row 
  # of the first column to the fifth column of the first row 
  
  # creating a place to store the reverse diagonal 
  reverse_diag <- 0
  for (i in 1:5){
    reverse_diag[i] <- card[i , 6-i]
  }
  
  # clause to check for a bingo, have an or to ensure that both 
  # directions of the diagonal are scanned by the function
  
  if (sum(reverse_diag) == 0 | sum(rev(reverse_diag)) == 0){
    print('Bingo! -- Reverse Diagonal!')
    return(card)
  } else if (length(card[card == 0]) >= 5){
    print('Keep playing!')
  } else {
    print('Keep playing!')
  }
  
}


```


And verifying. 

```{r}
result <- winner(card)
```

Now, we'll introduce a new function to play the game multiple times, with some 
introductory statements to codify the rules ahead of time, for ease of use. 

```{r}
full_game <- function(){
  card <- get_card()
  colSums(card)
  rowSums(card)
  length(card[card == 0])
  move_counter <- 0
  games_count <- 0
  diag_wins <- 0
  row_wins <- 0
  col_wins <- 0
  (result <- winner(card))
  #while (times_played < g){
    while (grepl('playing!', result) == TRUE){
      n <- sample(1:75 , 1)
      card <- marked(n, card)
      result <- winner(card)
      #return(card)
      #return(result)
      move_counter <- move_counter + 1
      # maybe something to just print the first result 
    }
    result
    if (grepl('playing!', result) == FALSE){
      print(length(result))
      print(result[1])
      print(paste('Number of Moves:', move_counter))
      times_played <- times_played + 1
      print(paste('Number of Games:', times_played))
      # start storing win types here 
      # something with this control flow is off 
      if (grepl('Diagonal', result[1]) == TRUE){
        diag_wins <- diag_wins + 1
      } else if (grepl('Rows!' , result[1]) == TRUE){
        row_wins <- row_wins + 1
      } else if (grepl('Columns!', result[1]) == TRUE){
        col_wins <- col_wins + 1
      }
    }
    wins_totals <- c(diag_wins , row_wins , col_wins)
    wins_totals
    
    for (i in 1:3){
      print(wins_totals[i])
    }
    
    games_count <- games_count + 1
    print(paste(
      paste('Diagonal:' , diag_wins) , 
      paste('Row:' , row_wins) , 
      paste('Col:' , col_wins) , 
      paste('Games Count:' , games_count)
    ))
    
    
    #print(paste('Diagonal:' , diag_wins))
    #print(paste('Row:' , row_wins))
    #print(paste('Col:' , col_wins))
    #games_count <- games_count + 1
    #print(paste('Games Count:', games_count , 
    #            ''))
  }
```

Here are some ideas for a multi-game framework and how we might simulate
the bingo game multiple times, commented out. 

```{r}

#for (i in 1:2){
#  test <- full_game()
#  print(test)
#}
# another function for the actual multi_game 

# maybe just do a for loop 
#results <- 0
#for (i in 1:3){
#  results <- full_game()
#  print(results)
#}



#multi_game <- function(n){
#  times_played <- 0
#  while (times_played < n){
#    full_game()
#    times_played <- times_played + 1
#  }
#}

#multi_game(2)

```




